Cover letter for the original IJRR submission

### 2017-10-02 reviews_tro/reviewer_response.tex

\done

\quot{
The first part of App. B is referred to within this Appendix as “new minor contributions derived from previous work”. These aren’t so novel, perhaps, and what is more useful to the reader is a better intuition for when to use each within the MAIN BODY of the paper, where they are originally described/associated to particular simulations.
}

\as{Agreed. We removed the incriminated text and provided an intuition of the heuristic in the text, as mentioned above.}\done

\quot{github link seems to be dead? (The link on p. 15 is OK\dots should they be the same?)}

\as{We are suprised by this comment because after testing it, the link seems to be valid. In any case, in accordance with reviewer 6, we changed all the links source codes to provide a SHA link that will bring the user to the accurate revision of the projects considered.}
\qt{(Section VII) Both HPP and our planner can be simply downloaded and compiled by following the instructions on
\url{https://humanoid-path-planner.github.io/hpp-doc/download.html?branch=rbprm}. \\
(Appendix C) To address the other scenarios, we propose a new implementation, entirely open source (\url{https://github.com/stonneau/python_sandbox/releases/tag/tro_paper} ), which can be integrated directly in our motion planner software. }\done

\quot{p.5, on right, ``less than epsilon'' between end effector and world - but doesn’t the orientation of the end effector matter tremendously here, too?}
\as{
The orientation of the end effector matters, but is not problematic regarding the way we generate contacts.
Using an epsilon is an effective way to remove many unrelevant configurations, and
we believe that more complex solutions aiming at taking into account the orientation would end up being counter productive.
 Indeed, using only 3D distance, we will end up with a few configurations
that can't be projected correctly, and will end up being discarded. But the majority of irrelevant configurations would already have been rejected with
the efficient 3D intersection query. Conversely, introducing a 6D octree rather than the current 3D one would make the problem much more complex: instead
of using well known 3D operations to retrieve our candidate configurations, we would have to implement our own, new solutions. The added cost of doing this is probably 
much higher than the computational cost of discarding a few configurations that could not be projected.  } \done

### 2017-10-02 body/intro.tex
finished pdf

Most of the papers that followed the work of Bretl have explored alternative formulations to handle the combinatorics. Two main directions have been explored. \textbf{On the one hand, local optimization of both the root trajectory \Pa and the contact positions $\mathcal{P}_2$} has been used, to trade the combinatorial of the complete problem for a differential complexity, at the cost of local convergence \citep{1631739}. A complete example of the potential offered by such approaches was proposed by \cite{Mordatch:2012:DCB:2185520.2185539} and successfully applied to a real robot \citep{mordatch2015}. To get reasonable computation times, the method uses a simplified dynamic model for the avatar. Still, the method is far from real-time  (about 1 minute of computation for 20 contacts).  A similar approach has been considered for manipulation by \cite{gabicciniisrr15}. Deits and Tedrake proposed to solve contact planning globally as a mixed-integer problem, but only cyclic, bipedal locomotion is considered, and equilibrium is not considered~\cite{DBLP:conf/humanoids/DeitsT14}. 
Dai et al.~\cite{dai2014whole} extended the work of Posa et al.~\cite{Posa:2014:DMT:2568343.2568352} to discover the contact sequence for landing motions, but need to specify
the contacts manually for complex interactions.
In addition to the limits of the current implementations, optimization-based approaches only converge locally.

>>>>

Most of the papers that followed the work of Bretl have explored alternative formulations to handle the combinatorics. Two main directions have been explored. \textbf{On the one hand, local optimization of both the root trajectory \Pa and the contact positions $\mathcal{P}_2$} has been used, to trade the combinatorial aspect of the complete problem for a differential complexity, at the cost of local convergence \citep{1631739}. A complete example of the potential offered by such approaches was proposed by \cite{Mordatch:2012:DCB:2185520.2185539} and successfully applied to a real robot \citep{mordatch2015}. To get reasonable computation times, the method uses a simplified dynamic model for the avatar. Still, the method is far from real-time  (about 1 minute of computation for 20 contacts).  A similar approach has been considered for manipulation by \cite{gabicciniisrr15}. Deits and Tedrake proposed to solve contact planning globally as a mixed-integer problem, but only cyclic, bipedal locomotion is considered, and equilibrium is not considered~\cite{DBLP:conf/humanoids/DeitsT14}. 
Dai et al.~\cite{dai2014whole} extended the work of Posa et al.~\cite{Posa:2014:DMT:2568343.2568352} to discover the contact sequence for landing motions, but need to specify
the contacts manually for complex interactions.
In addition to the limits of the current implementations, optimization-based approaches only converge locally.

### 2017-10-02 body/results.tex
finished pdf

\subsection{Description of the scenarios}
In all the scenarios considered, the formulation of the problem is always the same:
a start and goal root configurations are provided as input (except in the stair climbing scenario where the start whole body configuration is given).
The framework computes the initial contact configuration, and outputs a sequence of contact configurations connecting it to the goal.
In each scenario we detail the contacts involved and the heuristics chosen (either $h_{\textrm{\it EFORT}}$, $h_{vel}$ or $h_{w}$, all of which are defined in the Appendix~\ref{sec:heuristics}).
 %~ and the constraints on the reachable workspaces (for instance in all the scenarios, the reachable workspaces of the legs of HRP-2 are always required to intersect with the environment). 
%~ The companion video presents the complete contact sequence obtained in all these scenarios.

% \subsubsection*{Truck egress (Figure~\ref{res_truck_pres} and Figure~\ref{res_truck_bd}) -- Humanoid and insectoid robots.}
%\subsubsection*{Truck egress -- Humanoid and insectoid robots (Figure~\ref{res_truck_bd}).}
\subsubsection{HRP-2 -- Steep staircase (Figure~\ref{fig:stair_robust})}

\begin{figure}
  \centering
  \includegraphics[width=1\linewidth]{figures/stair}
  \caption{
           HRP-2 in the steep stair climbing scenario. }
		   \label{fig:stair_robust}
\end{figure}

@@

\subsubsection{Choosing the heuristics} \label{sec:heuristichoices}
In our conference paper~\citep{tonneauisrr15}, the computed motions were generated using the EFORT heuristic.
EFORT is designed for tasks requiring to exert important forces (such as pushing / pulling / climbing). 
In locomotion tasks, such as the stair scenario, one issue with EFORT is that it tends to generate
configurations close to singularities (and joint limits). While this did not significantly impact
the generation of the plan, the resulting interpolation turned out to be harder.
For this reason, we prefer to use our manipulability-based heuristic for the legs of the robot, but we still
use EFORT for the arms, which results in fewer contact repositionings.

@@

\subsection{Performance analysis} \label{sec:perf}
To analyze performance, we ran the planner 1000 times for each scenario.
We measured the computation time spent in each part of the algorithm, and analyzed their success rate.

@@

\begin{table}
\centering
\begin{tabular}{ l | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{35pt} | >{\centering\arraybackslash}m{35pt} | c}
  &  Equilibrium success rate & Kinematic failure & Equilibrium failure \\
 \hline
   Steep stairs 	& 99.5\%  & 0.1\% 	& 0.4\% \\
   Standing up 		& 87.8\%  & 6.1\% 	& 6.1\% \\
   Car egress 		& 66.2\%  & 15.9\% 	& 17.9\% \\
   Rubble 			& 97.54\% & 0.16\% 	& 2.3\% \\
   Obstacle race 	& 92.4\%  & 0.15\% 	& 7.45\% \\
 \end{tabular}
\caption{Success rates obtained for the generation of static equilibrium contact configurations for each scenario, rounded to the first decimal. Column 1 indicates 
indicates the rate of contact generation that succeeded. In the cases where the generation fails, it can be
either a kinematic issue (column 2), or because no contact configuration led to a static equilibrium configuration (column 3). Note that a failure in the contact generation
is not equivalent to a failure of the contact planning algorithm.}
\label{tab:requestpercent}
\quad
\end{table}

@@

Table~\ref{tab:requestpercent} presents the rate of successful contact generation. Note that a failure in contact generation for a root configuration is not equivalent to a failure in the contact plan. It simply means that another limb was tested for contact generation for the same root configuration.
As expected, the more constrained scenario, the car egress, provides the less satisfying results, despite the high success rate of the planner.

These results confirm that our approach provides a satisfying compromise between completeness and efficiency, thus allowing to consider online planning
while controlling the robot. Indeed, when the contact planning fails, it fails rapidly. This allows us to rapidly re-plan with a reasonable chance of success.
The most efficient (and immediate) approach to obtain a valid contact plan as fast as possible would be to launch in parallel several instances of the planner (our current implementation is single-threaded) and to use any successful result as a plan for solver $\mathcal{P}_3$.

>>>>

\subsection{Description of the scenarios}
In all the scenarios considered, the formulation of the problem is always the same:
a start and goal root configuration are provided as input (except in the stair climbing scenario where the start whole body configuration is given).
The framework computes the initial contact configuration, and outputs a sequence of contact configurations connecting it to the goal.
In each scenario we detail the contacts involved and the heuristics chosen (either $h_{\textrm{\it EFORT}}$, $h_{vel}$ or $h_{w}$, all of which are defined in the Appendix~\ref{sec:heuristics}).
 %~ and the constraints on the reachable workspaces (for instance in all the scenarios, the reachable workspaces of the legs of HRP-2 are always required to intersect with the environment). 
%~ The companion video presents the complete contact sequence obtained in all these scenarios.

% \subsubsection*{Truck egress (Figure~\ref{res_truck_pres} and Figure~\ref{res_truck_bd}) -- Humanoid and insectoid robots.}
%\subsubsection*{Truck egress -- Humanoid and insectoid robots (Figure~\ref{res_truck_bd}).}
\subsubsection{HRP-2 -- Staircase with high steps (Figure~\ref{fig:stair_robust})}

\begin{figure}
  \centering
  \includegraphics[width=1\linewidth]{figures/stair}
  \caption{
           HRP-2 in the stair climbing scenario. }
		   \label{fig:stair_robust}
\end{figure}

@@

\subsubsection{Choosing the heuristics} \label{sec:heuristichoices}
In our conference paper~\citep{tonneauisrr15}, the computed motions were generated using the EFORT heuristic.
EFORT is designed for tasks requiring large magnitude contact forces (such as pushing / pulling / climbing). 
In locomotion tasks, such as the stair scenario, one issue with EFORT is that it tends to generate
configurations close to singularities (and joint limits). While this did not significantly impact
the generation of the plan, the resulting interpolation turned out to be harder.
For this reason, we prefer to use our manipulability-based heuristic for the legs of the robot, but we still
use EFORT for the arms, which results in fewer contact repositionings.

@@

\subsection{Performance analysis} \label{sec:perf}
To analyze performance, we ran the planner 1000 times for each scenario.
We measured the computation time spent in each part of the algorithm, and analyzed success rate.

@@

\begin{table}
\centering
\begin{tabular}{ l | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{35pt} | >{\centering\arraybackslash}m{35pt} | c}
  &  Equilibrium success rate & Kinematic failure & Equilibrium failure \\
 \hline
   High stairs 	& 99.5\%  & 0.1\% 	& 0.4\% \\
   Standing up 		& 87.8\%  & 6.1\% 	& 6.1\% \\
   Car egress 		& 66.2\%  & 15.9\% 	& 17.9\% \\
   Rubble 			& 97.54\% & 0.16\% 	& 2.3\% \\
   Obstacle race 	& 92.4\%  & 0.15\% 	& 7.45\% \\
 \end{tabular}
\caption{Success rates obtained for the generation of static equilibrium contact configurations for each scenario, rounded to the first decimal. Column 1 indicates 
indicates the rate of contact generation that succeeded. In the cases where the generation fails, it can be
either a kinematic issue (column 2), or because no contact configuration led to a static equilibrium configuration (column 3). Note that a failure in the contact generation
is not equivalent to a failure of the contact planning algorithm.}
\label{tab:requestpercent}
\quad
\end{table}

@@

Table~\ref{tab:requestpercent} presents the rate of successful contact generation. Note that a failure in contact generation for a root configuration is not equivalent to a failure in the contact plan. It simply means that another limb was tested for contact generation for the same root configuration.
As expected, a more constrained scenario such as the car egress provides less satisfying results, despite the high success rate of the planner.

These results confirm that our approach provides a satisfying compromise between completeness and efficiency, thus enabling online planning
while controlling the robot. Indeed, when the contact planning fails, it fails rapidly. This allows us to rapidly re-plan with a reasonable chance of success.
The most efficient (and immediate) approach to obtain a valid contact plan as fast as possible would be to launch in parallel several instances of the planner (our current implementation is single-threaded) and to use any successful result as a plan for solver $\mathcal{P}_3$.
